Tor's Usability for Censorship Circumvention

### 2015-12-11 paper/circumvention-experiment.tex
Fix typos

There are still many features of Tor that are left unevaluated through user
research --- such as advanced web tasks (e.g., accessing hidden services), the
configuration menu to connect to the Tor network, configuring automatic
updates, and identity/cookie management. Rather than selecting the features to
study in isolation, we decided to focus on an important use case of Tor browser: censorship circumvention. 
Internet censorship is pervasive across the world today \cite{faris2008measuring}. 
Although Tor was not designed to be a censorship circumvention tool, users started
using it as once, since Tor's role as a proxy to anonymize
people also worked to circumvent censorship. Now, circumventing censorship is a
common use case for Tor, and current censorship circumvention methods are
incorporated into Tor, making it a viable censorship circumvention tool even in the face
of adversaries. In fact, some countries unsuccessfully attempt to block Tor for this reason \cite{winter2012great}. 
To our knowledge, this is the first user study investigating the usability of Tor as a 
censorship circumvention tool.

@@

\item {\bfseries Intermediate censorship} 
(Representative of countries such as Tunsia.)
Certain domains are blocked. Censorship circumvention
tools such as Tor are blocked. Since all public Tor
relay nodes are blocked, the default option to ``connect'' to the Tor network
directly will fail. Any choice of a hard-coded bridge
or a valid non-public bridge will circumvent this censor.  
Additional correct proxy configuration is optional.

@@

\noindent {\bfseries Procedure}
The experiment begins with the participants being informed that they are in a
simulated censorship
environment, where some --- but not all --- websites are blocked. We
instruct them to visit a non-blocked website and a blocked website on a
``standard'' browser (one that is not used for censorship circumvention, such
as Chrome, Safari, or Firefox) to illustrate the situation.
Participants are told that their
role is that of a regular citizen who is trying to visit blocked websites
for non-critical purposes,
and for whom the chance of risk is minimal. This information is given
to achieve
consistency across participants: behaviors may change if the tasks are
critical or if there were great risks associated with configuration mistakes.

@@

After users complete the browsing tasks (or time runs out), we conduct an
interview, collecting qualitative feedback from
participants about their browsing experience as a whole
and the configuration interface specifically.
We ask questions such as
``What was the most challenging part of connecting?''
and
``What is one change you would recommend?''

@@

@@

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item {\bfseries Observation 1: Users lack a mental model of how censorship works.} For instance, our participants did not know the difference between a bridge and a proxy or a censored website versus censored Tor relays. We suspect that this is the result of an accurate mental model of how the Internet works.  It would be unreasonable to assume that users should know this information, but some of this may be required to make a correct decision in the current interface. 
\item {\bfseries Observation 2: The interface flow does not guide users down the ideal path.} Much of the real estate on bridge configuration is used for the ``custom bridges'' option, which is rarely used. The interface instructs users to check out Internet settings to see if they would require a proxy, but it turns out that people are not skilled at doing so. Because the interface redirects users to the proxy screen after failing to make a connection, most of our participants tried configuring a proxy after a connection had failed due to an assumption that the proxy was the reason for the failure. In reality, there was no proxy setup required for any simulated censorship environment in our experiment and participants should have configured a bridge. 
\item {\bfseries Observation 3: There is little constructive feedback during
    failure cases.} When a connection seems to not work, our participants were
    unsure whether they needed to wait longer, or if they should try the
    connection with the same configuration again. Both of these are viable
    strategies, depending on the situation. Although there are certain error
    messages which were helpful, most of the error messages were technical and
    not understood by our participants (Figure \ref{fig:error}). The interface
    does not give clear directions on what participants should do next if a
    connection fails. 
\end{itemize}

Although we cannot comment on whether the challenges mentioned above will be
exactly representative of the Tor user population, it is fair to claim that some
Tor users will encounter them while configuring Tor. Our observations are
grounded in user interaction--specifically user mental modes, UI flow, and
failure design--giving us high confidence that addressing these observations
will make a noticeable impact on the usability of the Tor configuration dialog
for censorship circumvention.

@@

A clear example of this is the ``Connect'' button on the first screen
\ref{fig:redesign-firstscreen}, which has
increased in size and become green, making it more prominent as the recommended
option. We believe that this will help the majority of Tor users. 
Although some of our participants in environments had chosen to connect
directly when they needed to configure,
the majority of censorship today can be circumvented with a direct connection. 
A similar aesthetic applies to the connect button seen after custom
configuration.

@@

A progress indicator at the top of the window indicates which configuration step
the user is on. It serves to clarify the independence of bridges and proxies,
which need to be configured separately. In the event of a failure, the user is
returned to the first step of the process -- the bridge configuration -- to
avoid misleading users into thinking the proxy settings are the culprit. Previously,
the user had to manually click to go backward in the dialog to configure bridges.
Even in the rare event that the proxy is the issue for the connection setting, the user
will be directed to the proxy page naturally through the flow of the UI. \\

@@

\noindent {\bfseries Looking Forward} We will help incorporate changes into the Tor browser. Since we have developer support, it is not expected that we implement our changes, but we will advise them on any changes that they make. We hope that our work is able to help people more easily access the free Internet.

>>>>

There are still many features of Tor that are left unevaluated through user
research --- such as advanced web tasks (e.g., accessing hidden services), the
configuration menu to connect to the Tor network, configuring automatic
updates, and identity/cookie management. Rather than selecting the features to
study in isolation, we decided to focus on an important use case of Tor browser: censorship circumvention. 
Internet censorship is pervasive across the world today \cite{faris2008measuring}. 
Although Tor was not designed to be a censorship circumvention tool, users started
using it as one, since Tor's role as a proxy to anonymize
people also worked to circumvent censorship. Now, circumventing censorship is a
common use case for Tor, and current censorship circumvention methods are
incorporated into Tor, making it a viable censorship circumvention tool even in the face
of adversaries. In fact, some countries unsuccessfully attempt to block Tor for this reason \cite{winter2012great}. 
To our knowledge, this is the first user study investigating the usability of Tor as a 
censorship circumvention tool.

@@

\item {\bfseries Intermediate censorship} 
(Representative of countries such as Tunisia.)
Certain domains are blocked. Censorship circumvention
tools such as Tor are blocked. Since all public Tor
relay nodes are blocked, the default option to ``connect'' to the Tor network
directly will fail. Any choice of a hard-coded bridge
or a valid non-public bridge will circumvent this censor.  
Additional correct proxy configuration is optional.

@@

\noindent {\bfseries Procedure}
The experiment begins with the participants being informed that they are in a
simulated censorship
environment, where some --- but not all --- websites are blocked. We
instruct them to visit a non-blocked website and a blocked website on a
``standard'' browser (one that is not used for censorship circumvention, such
as Chrome, Internet Explorer, or Firefox) to illustrate the situation.
Participants are told that their
role is that of a regular citizen who is trying to visit blocked websites
for non-critical purposes,
and for whom the chance of risk is minimal. This information is given
to achieve
consistency across participants: behaviors may change if the tasks are
critical or if there are great risks associated with configuration mistakes.

@@

After users completed the browsing tasks (or time ran out), we conducted an
interview, collecting qualitative feedback from
participants about their browsing experience as a whole
and the configuration interface specifically.
We ask questions such as
``What was the most challenging part of connecting?''
and
``What is one change you would recommend?''

@@

% XXX: expect merge conflict here because whitespace was (accidentally?) removed in newer changes

@@

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item {\bfseries Observation 1: Users lack a mental model of how censorship works.} For instance, our participants did not know the difference between a bridge and a proxy or a censored website versus censored Tor relays. We suspect that this is the result of an inaccurate mental model of how the Internet works.  It would be unreasonable to assume that users should know this information, but some of this may be required to make a correct decision in the current interface. 
\item {\bfseries Observation 2: The interface flow does not guide users down the ideal path.} Much of the real estate on bridge configuration is used for the ``custom bridges'' option, which is rarely used. The interface instructs users to check out Internet settings to see if they would require a proxy, but it turns out that people are not skilled at doing so. Because the interface redirects users to the proxy screen after failing to make a connection, most of our participants tried configuring a proxy after a connection had failed due to an assumption that the proxy was the reason for the failure. In reality, there was no proxy setup required for any simulated censorship environment in our experiment and participants should have configured a bridge. 
\item {\bfseries Observation 3: There is little constructive feedback during
    failure cases.} When a connection seems to not work, our participants were
    unsure whether they needed to wait longer, or if they should try the
    connection with the same configuration again. Both of these are viable
    strategies, depending on the situation. Although there are certain error
    messages which were helpful, most of the error messages were technical and
    not understood by our participants (Figure \ref{fig:error}). The interface
    does not give clear directions on what participants should do next if a
    connection fails. 
\end{itemize}

Although we cannot comment on whether the challenges mentioned above will be
exactly representative of the Tor user population, it is fair to claim that some
Tor users will encounter them while configuring Tor. Our observations are
grounded in user interaction -- specifically user mental modes, UI flow, and
failure design -- giving us high confidence that addressing these observations
will make a noticeable impact on the usability of the Tor configuration dialog
for censorship circumvention.

@@

A clear example of this is the ``Connect'' button on the first screen
(Figure \ref{fig:redesign-firstscreen}), which has
increased in size and become green, making it more prominent as the recommended
option. We believe that this will help the majority of Tor users. 
Although some of our participants in environments had chosen to connect
directly when they needed to configure,
the majority of censorship today can be circumvented with a direct connection. 
A similar aesthetic applies to the connect button seen after custom
configuration.

@@

A progress indicator at the top of the window indicates which configuration step
the user is on. It serves to clarify the independence of bridges and proxies,
which need to be configured separately

In the event of a failure, the user is
returned to the first step of the process -- the bridge configuration -- to
avoid misleading users into thinking the proxy settings are the culprit. Previously,
the user had to manually click to go backward in the dialog to configure bridges.
In the rare event that the proxy is the issue for the connection setting, the user
will be directed to the proxy page naturally through the flow of the UI. \\

@@

\noindent {\bfseries Looking Forward} We will help incorporate changes into the Tor Browser. Since we have developer support, it is not expected that we implement our changes, but we will advise them on any changes that they make. We hope that our work is able to help people more easily access the free Internet.
